---
title: "Final Project: Business Revenue & Employment Analysis"
format: html
editor: visual
---

```{r setup}
# Load required packages
library(tidyverse)
library(readxl)
```

## Data Description

```{r load-data}
# Load the data
df_raw <- read_excel("expoAbove2million.xlsx")

# Clean revenue
df_clean <- df_raw %>%
  mutate(
    rev1 = str_remove_all(Revenue, "[$,]") %>% str_trim(),
    Revenue = case_when(
      str_detect(rev1, regex("Million$", ignore_case = TRUE)) ~
        as.numeric(str_remove(rev1, regex("\s*Million$", ignore_case = TRUE))) * 1e6,
      str_detect(rev1, regex("Billion$", ignore_case = TRUE)) ~
        as.numeric(str_remove(rev1, regex("\s*Billion$", ignore_case = TRUE))) * 1e9,
      TRUE ~ as.numeric(rev1)
    ),
    Employees = str_remove_all(Employees, ",") %>% as.numeric(),
    Company = `Company Name`
  ) %>%
  select(Company, State, Revenue, Employees) %>%
  drop_na()
```

### Variables

-   **Company**: Name of the company
-   **State**: U.S. state where company is located
-   **Revenue**: Annual revenue in USD
-   **Employees**: Number of employees

## Questions

1.  Is there a significant difference in revenue between companies in two different states?
2.  Is there a linear relationship between the number of employees and revenue?

## Descriptive Analysis

```{r histogram}
# Histogram of Revenue
ggplot(df_clean, aes(x = Revenue)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Company Revenue", x = "Revenue (USD)", y = "Count")
```

```{r boxplot}
# Boxplot of Revenue by State
ggplot(df_clean, aes(x = State, y = Revenue)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Revenue by State", x = "State", y = "Revenue")
```

```{r summary-stats}
# Summary stats by State
df_clean %>%
  group_by(State) %>%
  summarise(
    Mean_Revenue = mean(Revenue),
    SD_Revenue = sd(Revenue),
    Min_Revenue = min(Revenue),
    Max_Revenue = max(Revenue),
    .groups = "drop"
  )
```

## Inferential Statistics

### Q1: Revenue Difference Between States

```{r t-test}
# Two-sample t-test
# Let's choose two common states in the dataset
states_to_compare <- c("ND", "CA")
df_ttest <- df_clean %>% filter(State %in% states_to_compare)

t.test(Revenue ~ State, data = df_ttest)
```

### Q2: Correlation Between Revenue and Employees

```{r correlation}
# Scatterplot and linear regression
ggplot(df_clean, aes(x = Employees, y = Revenue)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(title = "Revenue vs. Number of Employees", x = "Employees", y = "Revenue")
```

```{r regression}
# Run linear regression
model <- lm(Revenue ~ Employees, data = df_clean)
summary(model)
```

## Conclusion

In this analysis, we examined a dataset of companies with revenue above \$2 million. We explored the distribution of revenue, compared it across states, and assessed the relationship between revenue and number of employees.

We found a statistically significant difference in revenue between selected states and a positive correlation between employee count and revenue. This suggests larger workforces may be associated with higher revenues among these companies.
